Jahn-Teller effect with rigid octahedral rotations in perovskites

### 2023-09-16 manuscript/paper.tex
intro paragraphs... literatiure... SOJT effect, nick etc. our appraach...

\documentclass[a4paper,prb]{revtex4-1}  %,twocolumn
%\documentclass[preprint]{revtex4-1}

@@

>>>>

\documentclass[a4paper,prb,twocolumn]{revtex4-1}  %,twocolumn
%\documentclass[preprint]{revtex4-1}

@@

\com{
para1: 
perovskites .. imp\\
octa rot imp...

para2: rot understood in terms of a tolerance factor... relative sizes of ions...\\
nothing substantial exists in the literature to explore the cause/reasons ... rotations...\\
except only recently, nick paper,... hydrogen bonding etc...

para3:
JT, ...
}

Perovskites are
technologically important materials 
that exhibit a rich variety of electronic, magnetic and optical properties
due to the interplay between \az{[refine...]}
orbital, spin, correlations, charge and structure.

orbital ordering... 
orbital-structure interplay.... structure-property relationships...

 \com{https://www.frontiersin.org/articles/10.3389/fchem.2020.550625/full:\\
advanced functional materials including photocatalysis, electrocatalysts
 high-temperature superconductors, multiferroics, magnetic and magnetoresistive materials, materials for solar cell energy and photoluminescence
}

\com{https://www.perovskite-info.com/introduction:\\
Perovskite materials exhibit intriguing and unusual physical properties that have been extensively studied for both practical applications and theoretical modeling and the materials science and applications of perovskites have been a broad research area open to many revolutionary discoveries for new device concepts. Perovskite?s potential applications are varied and include uses in sensors and catalyst electrodes, certain types of fuel cells, solar cells, lasers, memory devices and spintronics applications.
}

Their crystal structures have a common formula unit ABX$_3$,
where 
%positive ions A and negative ions X form a cube and an octahedron around positive ion B.
A and B are positive ions while X are negative ions 
%whose six-fold coordination forms octahedra around B.
that form octahedra around  B.
In a cubic perovskite,
the octahedra are symmetric, 
but they can distort or simply rotate to lower the symmetry
and produce a tetragonal or orthorhombic structure.
While the tetragonal distortion of the octahedra is associated with
a (cooperative) Jahn-Teller effect
(i.e., the orbital degeneracy of the metal ion coordinated by the anionic octahedron
is lifted, which stabilises the distortion),
%[as it distorts the octahedral field at the central site that split its electronic states],
the rotation and tilting of the octahedra are almost always 
explained in terms of 
the relative sizes of the three ions
with the
Goldschmidt tolerance factor. 
%[probably because rigid rotation of the octahedra do not change its potential at the central site]
%We only find two exceptions.

%\az{terminology: tilting, rotation.... vague is ok here...}

Based on density functional theory (DFT) calculations,
a few recent works
~\cite{garcia-fernandezJPCL2010,CammarataJCP2014,nick2016,YoshidaPRL21}, 
however, 
suggest that the octahedral rotation and tilting is 
a second order Jahn-Teller effect
%\az{describe SOJT in a sentence: coupling to the low lying unoccupied states on the JT ion, or orbital hybridisation with other ions, coupling between bands...}
(i.e., the distortion is stabilised by a coupling between the ground and excited adiabatic potential energy surfaces).
% garcia-fernandezJPCL2010
Garcia-Fernandez et. al.
study
potassium fluoride perovskite family KMF$_3$ (M$=$Ca$^{2+}$ or a $3$d transition metal)
and found that the tilting angle
depends on the covalent bonding through the mixing of occupied and 
low-lying unoccupied orbitals of the transition metal~\cite{garcia-fernandezJPCL2010}.
% CammarataJCP2014
Cammarata and Rondinelli
find a negative correlation between metal-oxygen bond covalency 
and octahedral rotation in orthorhombic perovskites oxides~\cite{CammarataJCP2014}.
% Lee, nick2016
Lee et. al. find 
similar results for metal-iodine bonding in
lead iodide perovskites,
where the role of 
the hydrogen bonding 
in a hybrid perovskite with
organic ammonium ions at A sites
is also found to be crucial~\cite{nick2016}.
Yoshida et. al.
use DFT along with 
representation theory analysis 
and 
indicate that 
the rotations are induced by bonding interaction 
between the valence and conduction band Bloch states at different wavevectors
that are mapped onto each other on Brillouin zone folding
 for the associated lattice distortion~\cite{YoshidaPRL21}.

We take a completely different approach to this problem.
We go back to the basics
and see 
what is it that lifts 
 the orbital degeneracy in the classical JT effect?
It is a deformation of the crystal field at the transition metal (JT ion) site.
Does the same happen in the case of octahedral tilting and rotations?
While the potential of the octahedral anion cage does not change in case of 
rigid rotations for obvious reason,
the net crystal field 
does change indeed
because the rest of the lattice is not only rotated (that turns its non-spherical potential)
in a given octahedron's frame but also distorted.
So, in principle, there should be a (first order) JT effect at play.
Using this insight
and elements of
crystal field theory,
here in this manuscript
we explore this problem in detail and 
find that the magnitude of this JT effect is sizeable for typical systems.

%1-liner summaries:\\
calcu sumamry...\\
analysis summary\\
results summary\\
We analytical results... numerical results... full lattice...

we explore states and energies of JT ion as a function of rotation and tilting angles...

and find that the
magnitude of the effect depends on the relative 
charges on the two cations involved, A and B,
and is sizeable for typical systems.

\com{
What is it that lifts the degeneracy in the classical JT effect?
its a deformation of the crystal field at the transition metal/JT ion site.
Does a deformation happen in the case of octahedral tilting and rotations?
Ridig octahedral.... octaherdal field of the octahedral anions cage is not changed 
... but... AB cages and the rest of the lattice is rotated and deformed... so... of course the CF should deform...
So there are in principle clear signs that there should be a (first order) JT effect at play.
}

\az{
Here, we .... just like the classical JT effect, where octahedral potential deforms....
....

we consider Pbnm.... other tilting and rotation.. glazer notation...???
}

\rev{structure of the paper:\\
